Forecasting short-term hourly Emergency Departement arrivals

### 2021-10-28 paper/paper.Rmd
change admision to arrival and hospital to ED

---
title:  Forecasting hourly hospital admission
author:
  - name: Author1
    email: email1@example.com
    affiliation: University1
    footnote: 1
  - name: Jethro Browell
    email: jethro.browell@glasgow.ac.uk
    affiliation: School of Mathematics and Statistics, University of Glasgow, UK
    footnote: 2
  - name: Ivan Svetunkov
    email: i.svetunkov@lancaster.ac.uk
    affiliation: Centre for Marketing Analytics and Forecasting, Lancaster University, UK
    footnote: 2
address:
  - code: University1
    address: Cardiff business school, 3 Colum Drive, CF10 3EU, Cardiff
  - code: University2
    address: adress2
  - code: University3
    address: adress3
footnote:
  - code: 1
    text: "Corresponding Author"
  - code: 2
    text: "Equal contribution"

@@

Forecasting hospital admission is a critical input to inform staffing and scheduling in hospital services to meet the needs of patients. An accurate hospital admission forecast contributes to a better decision making process regarding the resources needed to provide services to the number and type of patients requiring hospital services. This is one of the best ways to optimize staffing utilization and related costs. Most current practice to optimise personnel scheduling follows the general approach originally presented in @vile2016time, which recommends that the following steps be taken to roster employees: (i) forecast demand; (ii) convert demand forecasts into staffing requirements; (iii) schedule shifts optimally; and (iv) assign employees to shifts. An accurate demand forecasting is crucial in Accident & Emergency and Ambulance Services in particular, to depict various courses of action that can result in massive savings in terms of patient lives [@rostami2020anticipating].

Unability to match the staff with the demand results in patient overcrowding in the system which is a serious problem that causes challenging situations on patient flow [@derlet2002overcrowding]. Also, it is related with increasing length of stay [@muhammet2015forecasting], low patient satisfaction, number of patients left the hospital without seen by staff, unexpected return visits to services, increasing health care costs, inaccuracy in electronic medical record, and reported waiting times without incurring last-minute expenses, such as overtime or supplemental staffing.

There exists a large literature on forecasting hospital admission - including emergency department and ambulance services - that use various methods to forecast annual[@tandberg1994time] monthly,[@chen2011long; @mai2015predicting]. daily[@rostami2020anticipating;@park2019144] and subdaily [@schweigler2009forecasting;@cheng2021forecasting] admission. In this paper, we only focus on forecasting hourly hospital admission. In comparison to lower frequency time series forecasting such as monthly quarterly and yearly admission, hourly forecasts are challenging because the noise caused by random variation may overshadow any pattern in the time series.
<!-- In this respect, forecasts of daily or monthly arrivals are likely easier but target decisions about staff allocation and the like, not the ongoing scheduling and rescheduling of how the available resources are divided among the patients in need of emergency services --> An accurate forecast of the admission demand by hour of the day enables managers to match staff to meet anticipated patients, reconfigure units and redeploy staff. This will have many advantages for both patients, staff and the quality of provided services.  Hourly forecasts are required to inform the short-term operational planning for the current and the upcoming shifts of the day. This involves the short-term decision making related to the execution of the delivery process for various health care services such as Ambulatory, Emergency, Surgical, Inpatient, Home and Residential. The combination of an hourly forecast demand, current staff being occupied (e.g incidents being attended), resource availability and waiting times at hospitals, provide information on the state of the unscheduled care system across the service. Having this full picture enables the delivery managers to focus on the areas that require intervention to enable the most effective delivery of the service to the patients. Moreover, there are often unplanned events, such as walk-in patients, extended consultation times, during  a day or a week. Staff capacities may be adjusted based on the predicted demand fluctuations by using part-time, on-call nurses, staff overtime and voluntary absenteeism. That may also require to adjust the scheduling.

There are few studies that look at forecasting hourly hospital admission using historical time series data and/or predictors such as patient characteristics , climate factors , holidays. These studies use multiple approaches including Exponential Smoothing [REF], ARIMA families [], ARCH [REF], Vector Autoregressive [REF],  TBATS [] and ANN[]. Hourly time series generally exhibit multiple seasonal cycles of different lengths such as hourly, daily, weekly and yearly. They may also express nonstationarity and their profile may change over time. Therefore, an appropriate forecasting model should take these features into account, to accurately predict hourly demand admissions. This is currently missing in the literature and we fill this gap by examining various forecasting models capable of considering these features. All these studies only generate the estimated future admission as a point forecast (a single number), which does not report any uncertainty around the admission. Reporting uncertainty is critical because the consequences of imperfect staffing are asymmetric. Therefore, probabilistic forecasts are necessary to make decisions that balance the cost associated with under and over staffing. This asymmetry arises because it is preferable to incur a small opportunity cost associated under utilised staff rather than lower service levels if staff levels are insufficient. The lack of probabilistic forecast is one of the main bottlenecks in the deployment of generated forecasts in the staffing and planning tools. In this paper, we produce and evaluate the whole forecast distribution of hourly hospital admission, using all forecasting models, which could be used as a risk management tool for planners and decision makers. Aditionally, datasets used in some of these studies are relatively small, short length (e.g. time period of 1 year), which make it challenging to report the forecast accuracy using robust approaches such as time series cross validation and such results might not be reliable. Moreover, all previous publications  covered in this paper fails reproducibility principles (Boylan, 2016).

@@

1. We produce probabilistic forecasts, in addition to the point estimation, which quantify uncertainties in future hospital admission, and compare different forecasting methods using a suite of established evaluation metrics;

2. We develop a novel approach to forecast day-ahead hourly hospital admission using Generalised Additive Models for Location Scale and Shape, which accounts for i) the bounded and non-Gaussian distribution of admissions, ii) multiple seasonalities, weather and holiday effects, and  iii) variation in forecast uncertainty;

@@

```{r summarylit}
li <- readxl::read_xlsx("../table/table.xlsx")
knitr::kable(li, booktabs = T, linesep = "",caption = "Summary of studies in hourly emergency care forecasting") %>% kableExtra::kable_styling(latex_options = c("scale_down","hold_position"), font_size =11) %>% column_spec(3,"6em") %>%
  column_spec(4,"4em") %>% column_spec(6,"15em")  %>% column_spec(7,"4em") 
```

@@

We develop a range of models for forecasting hospital admissions using a multi-year dataset from a hospital in the UK. In order to facilitate model development and testing, we divide the data into a training and test period. We use cross-validation to iteratively fit, evaluate and develop models using the training data, before settling on their optimal configuration and testing them on the as-yet unseen test data. Data from 2014-04-01 to 2018-02-28 are used for training, and from 2018-03-01 to 2019-02-28 for testing.

Forecast are issued every day at midnight and mid-day for lead-times from 0 to 48 hours-ahead on a rolling basis. In order to compare both parametric and non-parametric probabilistic forecasts, we evaluate quantiles of the predictive distribution at probability level 0.05, 0.10,... 0.95, as well as the predictive expectation.

In the remainder of this section we explore the data in more detail, describe the forecast evaluation metrics we will employ, and introduce a range of benchmark models for producing forecasts. The benchmarks serve as reference methods and include some simple methods as well as more sophisticated routines designed to work `out-of-the-box', such as Facebook's Prophet. The performance of these methods provides an important reference, and any proposed novel solution should offer material benefits to justify increased complexity.

## Data exploration {#data}

The data we use in this case study are from hospital XYZ, and comprises records of arrivals at Accident and Emergency at hourly granularity. The training data runs from 01/04/2014 to 28/02/2018, and the test data runs from 1/3/2018 to 28/2/2019.

@@

The first benchmark we consider is to simply use the empirical distribution of past observations as a probabilistic forecast for same hour-of-day and day-of-week that we are forecasting. This is often called `climatology'. However, as we observe a trend in the data used here, we re-calculate the climatology on a rolling basis using the most recent one-year of data in order to track the trend.

@@

If we assume that our predictive distribution follows a given parametric distribution, the forecasting task becomes ones of predicting the future values of that distribution's parameters. Generalised Additive Models for Location, Scale and Shape (GAMLSS) are distributional regression models where the parameters of the assumed distribution may be modelled as additive functions of explanatory variables. This provides a powerful and flexible framework for probabilistic forecasting, provided that a suitable distribution and additive structures can be found. In practice, this means experimenting with various distributions and evaluating their suitability using available training data. Here we use this framework to extend the Poisson regression introduced above, and explore alternative distirbutional assumptions.

Let $y_t$ be the number of attendances in time period $t$ and indicate with $F_t(y_t)$ its predictive cumulative probability distribution. In a distributional regression context, $F_t(y_t)$ is modelled via a parametric model, $F(y_t|\bm \theta_t)$, where $\bm \theta_t$ is an $m$-dimensional vector of parameters. In a GAMLSS framework @Rigby2005 the elements $j=1,...,m$ of $\bm \theta_t$ are modelled via

@@

As our data are counts, the natural starting point is the Poisson distribution, given by
\begin{equation}
F_t(y_t,\lambda_t) =  \frac{\Gamma(\lfloor y_t + 1  \rfloor,\lambda_t)}{\lfloor y_t \rfloor}
(\#eq:poissonreg)
\end{equation}
<!-- where we initially consider an additive model for $\lambda_t$ of the form -->
<!-- \begin{equation} -->
<!--   \log(\lambda_t) = \sum_{i=1}^7 \beta_i D_i(t) + \sum_{j=1}^7 D_j(t) f_j(H(t)) + t f_\text{Y}(Y(t)) + f_\text{Temp}(Y(t),C_t) \quad , -->
<!--  (\#eq:additivemodel) -->
<!-- \end{equation} -->
<!-- where the functions $H(t)$ and $Y(t)$ return the hour of the day (1--24) and day of the year (1-366) at time $t$, respectively, $D_i(t)$ is a dummy variable for day of the week at $t$, and $C_t$ is the temperature at time $t$. Analysis of this model fit on the training data revealed systematic biases which were identified as relating to public and school holidays. We therefore expand the model to include these effects as follows, -->
<!-- \begin{equation} -->
<!--   \log(\lambda_t) = \sum_{i=1}^10 \beta_i D^{+}_i(t) + \sum_{j=1}^10 D^{+}_j(t) f_j(H(t)) + \sum_{k=1}^2 \beta_{10+k} S_k(t) + \sum_{l=1}^2 S_l(t) f_l(H(t)) + t + f_\text{Y}(Y(t)) + f_\text{Temp}(Y(t),C_t) \quad , -->
<!--  (\#eq:additivemodel_v2) -->
<!-- \end{equation} -->
<!-- where $D^{+}_i(t)$ is a dummy variable for day type including Monday to Sunday, Public Holiday, Christmas Day and New Year's Day, and $S_k(t)$ is a dummy variable for School Holidays. -->
where we consider an additive model for $\lambda_t$ of the form
\begin{equation}
\log(\lambda_t) = \sum_{i=1}^10 \beta_i D^{+}_i(t) + \sum_{j=1}^10 D^{+}_j(t) f_j(H(t)) + \sum_{k=1}^2 \beta_{10+k} S_k(t) + \sum_{l=1}^2 S_l(t) f_l(H(t)) + t + f_\text{Y}(Y(t)) + f_\text{Temp}(Y(t),C_t) \quad ,
(\#eq:additivemodel_v2)
\end{equation}
where the functions $H(t)$ and $Y(t)$ return the hour of the day (1--24) and day of the year (1-366) at time $t$, respectively, where $D^{+}_i(t)$ is a dummy variable for day type including Monday to Sunday, Public Holiday, Christmas Day and New Year's Day, $S_k(t)$ is a dummy variable for School Holidays, and $C_t$ is the temperature at time $t$. Public and school holidays were included following experiments on the training data which revealed systematic biases associated with these periods.

However, experiments on the training data reveal that calibration of forecasts based on the Poisson distribution is poor, suggesting that the shape of the distribution is unsuitable for the present application. In particular, we observe that forecast uncertainty appears to vary depending on the time of day and possibly other explanatory variables. Therefore, we consider more flexible, two parameter distributions so that we are able to specify additive models for both location and scale parameters, specifically the truncated Normal distribution, with truncation at 0. The resulting density forecasts are given by
\begin{equation}
F_t(y_t,\mu_t,\sigma_t) =  \frac{\Phi\left( \frac{y_t-\mu_t}{\sigma_t} \right) - \Phi\left( \frac{-y_t}{\sigma_t} \right)}{1 - \Phi\left( \frac{-y_t}{\sigma_t} \right)}
(\#eq:truncatedn)
\end{equation}
with additive models
\begin{align*}
\mu_t = &  \sum_{i=1}^10 \beta_i D^{+}_i(t) + \sum_{j=1}^10 D^{+}_j(t) f_j(H(t)) + \sum_{k=1}^2 \beta_{10+k} S_k(t) + \sum_{l=1}^2 S_l(t) f_l(H(t)) + t + f_\text{Y}(Y(t)) + f_\text{Temp}(Y(t),C_t) \quad, \\
log(\sigma_t) & = \sum_{i=1}^10 D^{+}_i(t) f(H(t))
(\#eq:additivemodel_NOtr_v2)
\end{align*}
for the mean and variance parameters. We also performed experiments with the truncated $t$ distribution and negative binomial distribution, but these did not results in forecasts as well calibrated as the truncated normal.

@@

The objective of density forecasts it to be as sharp as possible while remaining reliable/calibrated [@Gneiting2007a]. A forecast is said to be sharp if the predictive distribution has a relatively small spread, indicating low uncertainty, which is valuable to decision makers provided the forecast is calibrated.

The Pinball Score is a strictly proper score used to evaluate quantile forecasts and is the discrete form of the Continuous Rank Probability Score. It rewards sharpness and penalises mis-calibration, so measures all-round performance, however, calibration should still be verified separately. Furthermore, the Pinball Score for an individual quantile matches the loss function minimised in quantile regression model estimation. The Pinball Score is given by

@@

Calibration, also called `reliability', is the property that forecast probabilities match the observed frequency of realisations. If a forecast is calibrated, then $20\%$ of observations should fall below the $\alpha=0.2$ predictive quantile (with some tolerance based on the finite sample size). This property is necessary for forecast probabilities to be used in quantitative decision-making. Calibration is typically evaluated visually using reliability diagrams, which plot the nominal coverage, $\alpha$, against observer frequency mean($\mathbb{1}(y_{t}\le q_{\alpha,t})$).

@@

## Model fits

@@

@@

### Truncated Normal GAMLSS Model

We can support the interpretation the models given in Equation \@ref(eq:additivemodel_NOtr_v2) after the parameters have been estimated by visualising their reposes to inputs. Figure \@ref(fig-NOtr-model) shows the effect of time of day and day-type on the level and spread of predicted arrivals.

```{r fig-NOtr-model}

load("NOtr2_model.Rda")

plot_data <- data.frame(Hour = 0:23)
for(i in c("Mon","Tue","Wed","Thu","Fri","Sat","Sun","Public Holiday","Christmas Day","New Years Day")){
  pef <- gamlss::getPEF(h2_gamlss2,term="clock_hour",parameter = "mu",fixed.at = list(dow3=i))
  plot_data[[i]] <- pef(0:23)
}

plot_data <- reshape::melt.data.frame(plot_data,id.vars="Hour",variable_name="Day Type")

plot_1 <- ggplot(plot_data,aes(x=Hour,y=value,color=`Day Type`)) + geom_smooth(se = F,span=0.5) +
  ylab("Effect [Expected Arrivals]") + theme_few()

# plot_1

plot_data2 <- data.frame(Hour = 0:23)
for(i in c("Mon","Tue","Wed","Thu","Fri","Sat","Sun","Public Holiday","Christmas Day","New Years Day")){
  pef <- gamlss::getPEF(h2_gamlss2,term="clock_hour",parameter = "sigma",fixed.at = list(dow3=i))
  plot_data2[[i]] <- pef(0:23)
}

plot_data2 <- reshape::melt.data.frame(plot_data2,id.vars="Hour",variable_name="Day Type")

plot_2 <- ggplot(plot_data2,aes(x=Hour,y=value,color=`Day Type`)) + geom_smooth(se = F,span=0.5) +
  ylab("Effect [Standard Deviation of Arrivals]") + theme_few()

#plot_2

grid.arrange(plot_1, plot_2, ncol=2)

@@

```{r fig-quantilebias, fig.cap= "Quantile ...", fig.align='center',out.width= "90%"}
#knitr::include_graphics(rep("../results/QuantileBias.png"))
## Should reduce the number of models shown here... it's really hard to tell them appart!

REL_nom <- data.table(Nominal=seq(0,1,by=0.05),
                      Empirical=seq(0,1,by=0.05),
                      `Quantile Bias`= 0,
                      Method="Nominal")

REL[,`Quantile Bias`:=Empirical-Nominal]

ggplot(data=REL[Horizon=="All" & Issue == "All",],aes(x=Nominal,y=`Quantile Bias`,group=Method,color=Method)) +
  geom_line(data=REL_nom,aes(x=Nominal,y=`Quantile Bias`), color="black",size=1.1,show.legend = F) +
  geom_line() + geom_point() +
  xlim(c(0,1)) + ylim(c(-0.2,0.2)) + #ggtitle("Quantile Bias") +
  theme_few() 
```

@@

The Pinball Score evaluates overall performance, penalising both poor calibration and lack of sharpness, and ranks the various models quite differently. NBI-2 and Poisson-2 achive the lowest Pinball Loss, but as we have seen, are poorly calibrated so not suitable for use in decision making. Of the well calibrated methods, NOtr-2 and Ttr-2 have the lowest Pinball Score, and provide a significant improvement over Benchmark 2 by this metric. However, we note that while the improvement in terms of Pinball Score appears small, these methods have substantially improved forecast calibration compared to the benchmarks, which are both poorly calibrated.

@@

Forecast performance by time of day...

>>>>

---
title:  Forecasting short-term hourly Emergency Departement arrivals
author:
  - name: Author1
    email: email1@example.com
    affiliation: University1
    footnote: 1
  - name: Jethro Browell
    email: jethro.browell@glasgow.ac.uk
    affiliation: School of Mathematics and Statistics, University of Glasgow, UK
    footnote: 2
  - name: Ivan Svetunkov
    email: i.svetunkov@lancaster.ac.uk
    affiliation: Centre for Marketing Analytics and Forecasting, Lancaster University, UK
    footnote: 2
address:
  - code: University1
    address: Cardiff business school, 3 Colum Drive, CF10 3EU, Cardiff
  - code: University2
    address: adress2
  - code: University3
    address: adress3
footnote:
  - code: 1
    text: "Corresponding Author"
  - code: 2
    text: "Equal contribution"

@@

Forecasting Emergency Department (ED) arrivals is a critical input to inform staffing and scheduling decisions to meet the needs of patients. An accurate ED demand forecast contributes to a better decision making process regarding the resources needed to provide services to the number and type of patients requiring hospital services. This is one of the best ways to optimize staffing utilization and related costs. Most current practice to optimise personnel scheduling follows the general approach originally presented in @vile2016time, which recommends that the following steps be taken to roster employees: (i) forecast demand; (ii) convert demand forecasts into staffing requirements; (iii) schedule shifts optimally; and (iv) assign employees to shifts. An accurate demand forecasting is crucial in ED services to depict various courses of action that can result in massive savings in terms of patient lives. Unability to match the staff with the demand results in patient overcrowding in the system which is a serious problem that causes challenging situations on patient flow [@derlet2002overcrowding]. Also, it is related with increasing length of stay [@muhammet2015forecasting], low patient satisfaction, number of patients left the hospital without seen by staff, unexpected return visits to services, increasing health care costs, inaccuracy in electronic medical record, and reported waiting times without incurring last-minute expenses, such as overtime or supplemental staffing [@rostami2020anticipating].

There exists a large literature on forecasting ED arrivals that use various methods to forecast annual[@tandberg1994time] monthly,[@chen2011long; @mai2015predicting]. daily[@rostami2020anticipating;@park2019144] and subdaily [@schweigler2009forecasting;@cheng2021forecasting] arrivals. In this paper, we only focus on forecasting hourly arrivals. In comparison to lower frequency time series forecasting such as monthly quarterly and yearly, hourly forecasts are challenging because the noise caused by random variation may overshadow any pattern in the time series.
<!-- In this respect, forecasts of daily or monthly arrivals are likely easier but target decisions about staff allocation and the like, not the ongoing scheduling and rescheduling of how the available resources are divided among the patients in need of emergency services --> An accurate demand forecast by hour of the day enables planners to match staff to meet anticipated patients, reconfigure units and redeploy staff. This will have many advantages for both patients, staff and the quality of provided services. Hourly forecasts are required to inform the short-term operational planning for the current and the upcoming shifts of the day. This involves the short-term decision making related to the execution of the delivery process in ED.
<!-- for various health care services such as Ambulatory, Emergency, Surgical, Inpatient, Home and Residential.  -->
The combination of an hourly forecast demand, current staff being occupied, resource availability and waiting times at ED, provide information on the state of the unscheduled care system across the service. Having this full picture enables the delivery managers to focus on the areas that require intervention to enable the most effective delivery of the service to the patients. Moreover, there are often unplanned events, such as walk-in patients, extended consultation times, during  a day or a week. Staff capacities may be adjusted based on the predicted demand fluctuations by using part-time, on-call nurses, staff overtime and voluntary absenteeism. That may also require to adjust the scheduling.

There are few studies that look at forecasting hourly arrivals in ED and other hospital services using historical time series data and/or predictors such as patient characteristics , climate factors , holidays. These studies use multiple approaches including Exponential Smoothing [REF], ARIMA families [], ARCH [REF], Vector Autoregressive [REF],  TBATS [] and ANN[]. Hourly time series generally exhibit multiple seasonal cycles of different lengths such as hourly, daily, weekly and yearly. They may also express nonstationarity and their profile may change over time. Therefore, an appropriate forecasting model should take these features into account, to accurately predict hourly demand admissions. This is currently missing in the literature and we fill this gap by examining various forecasting models capable of considering these features. All these studies only generate the estimated future admission as a point forecast (a single number), which does not report any uncertainty around the admission. Reporting uncertainty is critical because the consequences of imperfect staffing are asymmetric. Therefore, probabilistic forecasts are necessary to make decisions that balance the cost associated with under and over staffing. This asymmetry arises because it is preferable to incur a small opportunity cost associated under utilised staff rather than lower service levels if staff levels are insufficient. The lack of probabilistic forecast is one of the main bottlenecks in the deployment of generated forecasts in the staffing and planning tools. In this paper, we produce and evaluate the whole forecast distribution of hourly hospital admission, using all forecasting models, which could be used as a risk management tool for planners and decision makers. Aditionally, datasets used in some of these studies are relatively small, short length (e.g. time period of 1 year), which make it challenging to report the forecast accuracy using robust approaches such as time series cross validation and such results might not be reliable. Moreover, all previous publications  covered in this paper fails reproducibility principles.

@@

1. We develop a novel methodology to forecast short-term hourly hospital admission using the family of Generalised Additive Models that accounts for i) ..., ii) ... iii) , ... iv) (Jethro to complete)

2. We produce probabilistic forecast, in addition to the point estimation, that quantify uncertainties in future hospital admission and we evaluate its accuracy using ...;

@@

```{r summarylit}
li <- readxl::read_xlsx("../table/table.xlsx")
knitr::kable(li, booktabs = T, linesep = "",caption = "Summary of studies in hourly emergency care forecasting") %>% kableExtra::kable_styling(latex_options = c("scale_down","hold_position"), font_size =11) %>% column_spec(3,"6em") %>%
  column_spec(4,"4em") %>% column_spec(6,"15em")  %>% column_spec(7,"4em") %>%  kableExtra::landscape()
```

@@

## data {#data}

@@

Empirical distribution for same hour-of-day and day-of-week as target time.

@@

If we assume that our predictive distribution follows a given parametric distribution, the forecasting task becomes ones of predicting the future values of that distribution's parameters. Generalised Additive Models for Location, Scale and Shape (GAMLSS) are distributional regression models where the parameters of the assumed distribution may be modelled as additive functions of explanatory variables. This provides a powerful and flexible framework for probabilistic forecasting, provided that a suitable distribution and additive structures can be found. In practice, this means experimenting with various distributions and evaluating their suitability using available training data.
Let $y_t$ be the number of attendances in time period $t$ and indicate with $F_t(y_t)$ its predictive cumulative probability distribution. In a distributional regression context, $F_t(y_t)$ is modelled via a parametric model, $F(y_t|\bm \theta_t)$, where $\bm \theta_t$ is an $m$-dimensional vector of parameters. In a GAMLSS framework @Rigby2005 the elements $j=1,...,m$ of $\bm \theta_t$ are modelled via

@@

As our data are counts, the natural starting point is the Poisson distribution, given by

\begin{equation}
  F_t(y_t,\lambda_t) =  \frac{\Gamma(\lfloor y_t + 1  \rfloor,\lambda_t)}{\lfloor y_t \rfloor}
  (\#eq:poissonreg)
\end{equation}

where we consider an additive model for $\lambda_t$ of the form

\begin{equation}
  \log(\lambda_t) = \sum_{i=1}^7 \beta_i \delta(D_i(t)-i) + \sum_{j=1}^7 D_j(t) f_j(H(t)) + t f_\text{Y}(Y(t)) + f_\text{Temp}(Y(t),C_t) \quad .
 (\#eq:additivemodel)
\end{equation}

The functions $H(t)$, $D(t)$ and $Y(t)$ return the hour of the day (1--24), day of the week (1--7), and day of the year (1-366) at time $t$, respectively, and $C_t$ is the temperature at time $t$.

Truncated Normal...

\begin{equation}
  F_t(y_t,\mu_t,\sigma_t) =  \frac{\Phi\left( \frac{y_t-\mu_t}{\sigma_t} \right) - \Phi\left( \frac{-y_t}{\sigma_t} \right)}{1 - \Phi\left( \frac{-y_t}{\sigma_t} \right)}
 (\#eq:truncatedn)
\end{equation}

with

\begin{align*}
  \mu_t = &  \sum_{i=1}^7 \beta_i \delta(D_i(t)-i) + \sum_{j=1}^7 D_j(t) f_j(H(t)) + t f_\text{Y}(Y(t)) + f_\text{Temp}(Y(t),C_t) \quad, \\
  log(\sigma_t) & = f(H(t)) .
\end{align*}

Truncated $t$ distribution...

Negative binomial...

@@

The objective of density forecasts it to be as sharp as possible while remaining reliable/calibrated [@Gneiting2007a]. A forecast is said to be sharp if the predictive distribution has a relatively small spread, indicating low uncertainty, witch is valuable to decision makers provided the forecast is calibrated.

The Pinball Score is a strictly proper score used to evaluate quantile forecasts and is the discrete form of the Continuous Rank Probability Score. It rewards sharpness and penalises mis-calibration, so measures all-round performance, however, calibration should still be verified separately. Furthermore, 
The Pinball Score for an individual quantile matches the loss function minimised in quantile regression model estimation. The Pinball Score is given by

@@

Calibration, also called `reliability`, is the property that forecast probabilities match the observed frequency of realisations. If a forecast is calibrated, then $20\%$ of observations should fall below the $\alpha=0.2$ predictive quantile (with some tolerance based on the finite sample size). This property is necessary for forecast probabilities to be used in quantitative decision-making. Calibration is typically evaluated visually using reliability diagrams, which plot the nominal coverage, $\alpha$, against observer frequency mean($\mathbb{1}(y_{t}\le q_{\alpha,t})$).

@@

@@

## Case study

The data is portioned into training and test data, with all model development and hyper-parameter tuning performed using training data only. Data from 2014-04-01 to 2018-02-28 are used for training, and from 2018-03-01 to 2019-02-28 for testing.

@@

@@

# ```{r fig-quantilebias, fig.cap= "Quantile ...", fig.align='center',out.width= "90%"}
# #knitr::include_graphics(rep("../results/QuantileBias.png"))
# ## Should reduce the number of models shown here... it's really hard to tell them appart!
# 
# REL_nom <- data.table(Nominal=seq(0,1,by=0.05),
#                       Empirical=seq(0,1,by=0.05),
#                       `Quantile Bias`= 0,
#                       Method="Nominal")
# 
# REL[,`Quantile Bias`:=Empirical-Nominal]
# 
# ggplot(data=REL[Horizon=="All" & Issue == "All",],aes(x=Nominal,y=`Quantile Bias`,group=Method,color=Method)) +
#   geom_line(data=REL_nom,aes(x=Nominal,y=`Quantile Bias`), color="black",size=1.1,show.legend = F) +
#   geom_line() + geom_point() +
#   xlim(c(0,1)) + ylim(c(-0.2,0.2)) + #ggtitle("Quantile Bias") +
#   theme_few() 
# ```

@@

@@
